Evolution of the AGN UV luminosity function

### 2017-07-10 Paper/paper.tex
Update text

At $4.7\le z<5.1$ we combined several recent surveys, accounting for
sample overlap and updated selection functions. At the bright end of
the QLF we used the 71 $z<5.1$ quasars from the SDSS+WISE survey
\citep{2016ApJ...829...33Y} together with its selection function.
%Although \citet{yang} are sensitive to $z>5.1$ quasars, we limit
%their sample to $z<5.1$ to limit differential evolution of the QLF
%across the redshift range.
To avoid double-counting quasars contained both in
\citet{2016ApJ...829...33Y} and the SDSS DR7 sample from
\citet{2013ApJ...768..105M}, we used the latter only at
$M_{1450}>-26.72$ (in our adopted cosmology), yielding 94 additional
$4.7\le z<5.1$ quasars selected in 6248\,deg$^2$. We used the updated
$z\sim 5$ SDSS DR7 selection function \citep{2013ApJ...768..105M}
instead of the one from \citet{2006AJ....131.2766R}.
%\footnote{We note that the major difference in the selection
%functions must be due to different input parameters, since the single
%color selection criterion in \citet{mcgreer} selects the vast
%majority of the SDSS DR7 quasars.}.
To these two bright-end samples we added the faint-end sample from the
\citet{2013ApJ...768..105M} SDSS Stripe~82 survey (52 uniformly
selected $4.7\le z<5.1$ quasars in 235\,deg$^2$) and two quasars from
\citet{2011ApJ...728L..26G}, adopting the respective selection
functions. We did not consider the limit on the $z\sim 5$ QLF by
\citet{2012ApJ...756..160I} due to systematic errors in their
selection function\footnote{\citet{2012ApJ...756..160I} underestimated
  the dispersion in rest-frame UV quasar colours with respect to SDSS
  at all redshifts (their Figure~4). Contrary to their claim,
  photometric errors have a small effect on the colour distribution of
  SDSS quasars given the statistical errors of $<0.03$\,mag in $gri$
  for 90 per cent of the SDSS DR7 bright quasar sample ($i<19.1$) and
  a relative calibration error of $\sim 1$ per cent
  \citep{2008ApJ...674.1217P}.}.

@@

For a joint fit of the QLF it is necessary to homogenise the different
survey samples in absolute magnitude, and to convert their selection
functions to the same absolute magnitude system. For the analysis of
the quasar UV emissivity and to be consistent with published work at
$z>3$ we chose to convert all samples and selection functions to the
absolute AB magnitude at a rest frame wavelength $\lambda=1450$\,\AA
\begin{equation}\label{eq:absmag}
M_{1450}\left(z\right) = m-5\log{\left(\frac{d_L\left(z\right)}{\mathrm{Mpc}}\right)}-25-K_{m,1450}\left(z\right),
\end{equation}
with the luminosity distance $d_L\left(z\right)$ to a quasar at
redshift $z$, the apparent magnitude $m$ in a filter used in the
survey, and the bandpass correction $K_{m,1450}\left(z\right)$
\citep{1956AJ.....61...97H, 1968ApJ...154...21O, 2000A&A...353..861W,
  2002astro.ph.10394H}. For the bandpass correction we used a
combination of the \citet{2015MNRAS.449.4204L} quasar SED at
$\lambda<2500$\,\AA, and the \citet{2001AJ....122..549V} quasar
composite spectrum at longer wavelengths to cover the lowest
redshifts.  The samples from SDSS and BOSS are defined in the SDSS $i$
band, while 2SLAQ is defined in the $g$ band. At $z>4.7$ we adopted
the SDSS $z$ band magnitude (in the following denoted $z_\mathrm{AB}$)
for SDSS DR7 quasars to avoid additional corrections due to the
Ly$\alpha$ forest. Figure~XX shows our bandpass corrections for SDSS,
BOSS and 2SLAQ as a function of redshift. While the
\citet{2015MNRAS.449.4204L} SED is for bright ($M_{1450}\simeq -27.2$)
quasars, UV composite spectra including fainter quasars
\citep{2002ApJ...565..773T, 2012ApJ...752..162S, 2014ApJ...794...75S}
give similar values, such that our bandpass corrections likely remain
applicable at $M_{1450}\la -24$. \textbf{What about even fainter AGN?}

For the $z<2.2$ sample we corrected the SDSS $i$ and 2SLAQ $g$ band
magnitudes for host galaxy contamination following
\citet{2009MNRAS.392...19C}. Considering the different magnitude
limits of 2SLAQ and SDSS, the modelled host galaxy contamination is
small for $z>0.5$ quasars ($<0.1$\,mag in $g$, $<0.2$\,mag in $i$),
and is negligible at $z>0.8$.  In case the band defining the magnitude
limit of the survey undesirably overlaps with the Ly$\alpha$ forest
\citep{2010ApJ...710.1498G, 2011ApJ...728L..26G, 2013ApJ...768..105M}
we adopted their respective bandpass corrections to $M_{1450}$. In
particular, for the \citet{2010ApJ...710.1498G, 2011ApJ...728L..26G}
sample we recomputed $M_{1450}$ from the $R$ band photometry to be
consistent with the selection function defined in $R$, and to avoid
uncertainties in their spectrophotometry due to incomplete spectral
coverage. For the remaining high-redshift surveys reporting $M_{1450}$
obtained by various methods \citep{2010AJ....139..906W,
  2015ApJ...798...28K, 2015AA...578A..83G, 2016ApJ...829...33Y,
  2016ApJ...833..222J} we did not re-compute $M_{1450}$, but applied
appropriate shifts to correct to our adopted cosmology.

@@

For 2SLAQ we corrected for magnitude-dependent spectroscopic coverage
in the two survey areas \citep[$f_\mathrm{c,NGP}\left(g\right)$ and
  $f_\mathrm{c,SGP}\left(g\right)$; Figure~4
  in][]{2009MNRAS.392...19C} and spectroscopic redshift success
\citep[$f_\mathrm{s,2SLAQ}\left(g\right)$; Figure~6b
  in][]{2009MNRAS.392...19C} by multiplying them with the photometric
selection function, resulting in two area-specific 2SLAQ selection
functions
$f_\mathrm{NGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,NGP}f_\mathrm{s,2SLAQ}$
and
$f_\mathrm{SGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,SGP}f_\mathrm{s,2SLAQ}$
that are relevant for the QLF.  The $z<4.7$ SDSS photometric selection
function was modified to include known imaging incompleteness to
$f_{\mathrm{SDSS},z<4.7}=0.95f_{\mathrm{p,SDSS},z<4.7}$
\citep{2006AJ....131.2766R}. The BOSS colour-selected sample contains
quasars with $f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}\ge 0.85$
\citep{2013ApJ...773...14R}, and we adopted
$f_\mathrm{BOSS}=\overline{f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}}f_\mathrm{p,BOSS}=0.962f_\mathrm{p,BOSS}$. \citet{2010ApJ...710.1498G}
presented two area-specific photometric selection functions due to
different filters employed, and more follow-up spectroscopy was
reported in \citet{2011ApJ...728L..26G}. We accounted for remaining
spectroscopic incompleteness at $R>23$, yielding the final selection
functions $f_\mathrm{NDWFS}$ and $f_\mathrm{DLS}$. The updated $z\sim
5$ SDSS photometric selection function \citep{2013ApJ...768..105M} was
modified to include imaging and spectroscopic incompleteness, yielding
$f_{\mathrm{SDSS},z\sim 5}=0.95^2f_{\mathrm{p,SDSS},z\sim 5}$. In the
deeper $z\sim 5$ SDSS Stripe~82 survey the spectroscopic
incompleteness is larger and magnitude-dependent \citep[Figure~14
  in][]{2013ApJ...768..105M}, resulting in $f_{\mathrm{S82},z\sim
  5}=0.95f_{\mathrm{s,S82},z\sim
  5}\left(i\right)f_{\mathrm{p,S82},z\sim 5}$. Likewise, imaging and
magnitude-dependent spectroscopic incompleteness was factored into the
\citet{2016ApJ...829...33Y} photometric selection function (their
Figures~5 and 7), resulting in
$f_\mathrm{SDSS+WISE}=0.97f_\mathrm{s,SDSS+WISE}\left(z_\mathrm{AB}\right)f_\mathrm{p,SDSS+WISE}$. Finally,
we obtained a rough estimate of the \citet{2015AA...578A..83G}
selection function by taking
$f_\mathrm{GOODS-S}=\phi_\mathrm{obs}/\phi_\mathrm{corr}$ (see their
Table~3).

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{emissivity_18.pdf}
    \includegraphics[width=0.5\textwidth]{emissivity_20.pdf}
    \end{tabular}
  \end{center}
  \caption{LyC emissivity of AGN assuming 100\% escape fraction.
    Model luminosity functions are integrated down to $M_{1450}=-18$
    in the left panel and $-20$ in the right panel.}
  \label{fig:emissivity}
\end{figure*}

>>>>

At $4.7\le z<5.1$ we combined several recent surveys, accounting for sample overlap and updated selection functions. At the bright end of the QLF we used the 71 $z<5.1$ quasars from the SDSS+WISE survey \citep{2016ApJ...829...33Y} together with its selection function.
%Although \citet{yang} are sensitive to $z>5.1$ quasars, we limit their sample to $z<5.1$ to limit differential evolution of the QLF across the redshift range.
To avoid double-counting quasars contained both in \citet{2016ApJ...829...33Y} and the SDSS DR7 sample from \citet{2013ApJ...768..105M}, we used the latter only at $M_{1450}>-26.72$ (in our adopted cosmology), yielding 94 additional $4.7\le z<5.1$ quasars selected in 6248\,deg$^2$. We used the updated $z\sim 5$ SDSS DR7 selection function \citep{2013ApJ...768..105M} instead of the one from \citet{2006AJ....131.2766R}.
%\footnote{We note that the major difference in the selection functions must be due to different input parameters, since the single color selection criterion in \citet{mcgreer} selects the vast majority of the SDSS DR7 quasars.}.
To these two bright-end samples we added the faint-end sample from the \citet{2013ApJ...768..105M} SDSS Stripe~82 survey (52 uniformly selected $4.7\le z<5.1$ quasars in 235\,deg$^2$) and two quasars from \citet{2011ApJ...728L..26G}, adopting the respective selection functions. We did not consider the limit on the $z\sim 5$ QLF by \citet{2012ApJ...756..160I} due to systematic errors in their selection function\footnote{\citet{2012ApJ...756..160I} underestimated the dispersion in rest-frame UV quasar colours with respect to SDSS at all redshifts (their Figure~4). Contrary to their claim, photometric errors have a small effect on the colour distribution of SDSS quasars given the statistical errors of $<0.03$\,mag in $gri$ for 90 per cent of the SDSS DR7 bright quasar sample ($i<19.1$) and a relative calibration error of $\sim 1$ per cent \citep{2008ApJ...674.1217P}.}.

@@

\begin{figure}
    \includegraphics[width=\columnwidth]{kcorr.pdf}
  \caption{Bandpass corrections $K_{m,1450}$ from a broadband magnitude $m=\{g,i,z_\mathrm{AB}\}$
           to the monochromatic AB magnitude at 1450\,\AA\ as a function of redshift $z$ for the \citet{2015MNRAS.449.4204L}
           quasar SED used in this work, and for two quasar composite spectra \citep{2001AJ....122..549V, 2002ApJ...565..773T}.
           The redshift range has been restricted to exclude the Ly$\alpha$ forest and to account for the different
           rest frame wavelength coverage of the spectra.}
  \label{fig:kcorr}
\end{figure}

For a joint fit of the QLF it is necessary to homogenise the different survey samples in absolute magnitude, and to convert their selection functions to the same absolute magnitude system. For the analysis of the quasar UV emissivity and to be consistent with published work at $z>3$ we chose to convert all samples and selection functions to the absolute AB magnitude at a rest frame wavelength $\lambda=1450$\,\AA
\begin{equation}\label{eq:absmag}
M_{1450}\left(z\right) = m-5\log{\left(\frac{d_L\left(z\right)}{\mathrm{Mpc}}\right)}-25-K_{m,1450}\left(z\right),
\end{equation}
with the luminosity distance $d_L\left(z\right)$ to a quasar at redshift $z$, the apparent magnitude $m$ in a filter used in the survey, and the bandpass correction $K_{m,1450}\left(z\right)$ \citep{1956AJ.....61...97H, 1968ApJ...154...21O, 2000A&A...353..861W, 2002astro.ph.10394H}. For the bandpass correction we used a combination of the \citet{2015MNRAS.449.4204L} quasar SED at $\lambda<2500$\,\AA, and the \citet{2001AJ....122..549V} quasar composite spectrum at longer wavelengths to cover the lowest redshifts. 
The samples from SDSS and BOSS are defined in the SDSS $i$ band, while 2SLAQ is defined in the $g$ band. At $z>4.7$ we adopted the SDSS $z$ band magnitude (in the following denoted $z_\mathrm{AB}$) for SDSS DR7 quasars to avoid additional corrections due to the Ly$\alpha$ forest. Figure~\ref{fig:kcorr} shows our bandpass corrections for SDSS, BOSS and 2SLAQ as a function of redshift. We ignored the luminosity dependence of the bandpass correction due to the known anticorrelation of emission line equivalent width and luminosity \citep{1977ApJ...214..679B}.  While the \citet{2015MNRAS.449.4204L} SED is for luminous ($M_{1450}\simeq -27.2$) quasars, UV composite spectra including fainter quasars \citep{2002ApJ...565..773T, 2012ApJ...752..162S, 2014ApJ...794...75S} give similar values, such that our bandpass corrections remain applicable at $M_{1450}\la -24$. Empirical luminosity-dependent bandpass corrections show a $\la 0.2$\,mag variation over $\sim 4$ orders of absolute magnitude depending on redshift and the filter, and with a $\sim 0.2$\,mag intrinsic scatter due to individual SED variations \citep{2013ApJ...773...14R, 2013ApJ...768..105M, 2013A&A...551A..29P}.

For the $z<2.2$ sample we corrected the SDSS $i$ and 2SLAQ $g$ band magnitudes for host galaxy contamination following \citet{2009MNRAS.392...19C}. Considering the different magnitude limits of 2SLAQ and SDSS, the modelled host galaxy contamination is small for $z>0.5$ quasars ($<0.1$\,mag in $g$, $<0.2$\,mag in $i$), and is negligible at $z>0.8$.
In case the band defining the magnitude limit of the survey undesirably overlaps with the Ly$\alpha$ forest \citep{2010ApJ...710.1498G, 2011ApJ...728L..26G, 2013ApJ...768..105M} we adopted their respective bandpass corrections to $M_{1450}$. In particular, for the \citet{2010ApJ...710.1498G, 2011ApJ...728L..26G} sample we recomputed $M_{1450}$ from the $R$ band photometry to be consistent with the selection function defined in $R$, and to avoid uncertainties in their spectrophotometry due to incomplete spectral coverage. Since the \citet{2010ApJ...710.1498G, 2011ApJ...728L..26G} $R$ band traces the rest frame UV, we assumed negligible host galaxy contamination for their faint quasars. For the remaining high-redshift surveys reporting $M_{1450}$ obtained by various methods \citep{2010AJ....139..906W, 2015ApJ...798...28K, 2015AA...578A..83G, 2016ApJ...829...33Y, 2016ApJ...833..222J} we did not re-compute $M_{1450}$, but applied appropriate shifts to correct to our adopted cosmology.

@@

For 2SLAQ we corrected for magnitude-dependent spectroscopic coverage in the two survey areas \citep[$f_\mathrm{c,NGP}\left(g\right)$ and $f_\mathrm{c,SGP}\left(g\right)$; Figure~4 in][]{2009MNRAS.392...19C} and spectroscopic redshift success \citep[$f_\mathrm{s,2SLAQ}\left(g\right)$; Figure~6b in][]{2009MNRAS.392...19C} by multiplying them with the photometric selection function, resulting in two area-specific 2SLAQ selection functions $f_\mathrm{NGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,NGP}f_\mathrm{s,2SLAQ}$ and $f_\mathrm{SGP}\left(M_{1450},z\right)=f_\mathrm{p,2SLAQ}f_\mathrm{c,SGP}f_\mathrm{s,2SLAQ}$ that are relevant for the QLF.
The $z<4.7$ SDSS photometric selection function was modified to include known imaging incompleteness to $f_{\mathrm{SDSS},z<4.7}=0.95f_{\mathrm{p,SDSS},z<4.7}$ \citep{2006AJ....131.2766R}. The BOSS colour-selected sample contains quasars with $f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}\ge 0.85$ \citep{2013ApJ...773...14R}, and we adopted $f_\mathrm{BOSS}=\overline{f_\mathrm{c,BOSS}f_\mathrm{s,BOSS}}f_\mathrm{p,BOSS}=0.962f_\mathrm{p,BOSS}$. \citet{2010ApJ...710.1498G} presented two area-specific photometric selection functions due to different filters employed, and more follow-up spectroscopy was reported in \citet{2011ApJ...728L..26G}. We accounted for remaining spectroscopic incompleteness at $R>23$, yielding the final selection functions $f_\mathrm{NDWFS}$ and $f_\mathrm{DLS}$. The updated $z\sim 5$ SDSS photometric selection function \citep{2013ApJ...768..105M} was modified to include imaging and spectroscopic incompleteness, yielding $f_{\mathrm{SDSS},z\sim 5}=0.95^2f_{\mathrm{p,SDSS},z\sim 5}$. In the deeper $z\sim 5$ SDSS Stripe~82 survey the spectroscopic incompleteness is larger and magnitude-dependent \citep[Figure~14 in][]{2013ApJ...768..105M}, resulting in $f_{\mathrm{S82},z\sim 5}=0.95f_{\mathrm{s,S82},z\sim 5}\left(i\right)f_{\mathrm{p,S82},z\sim 5}$. Likewise, imaging and magnitude-dependent spectroscopic incompleteness was factored into the \citet{2016ApJ...829...33Y} photometric selection function (their Figures~5 and 7), resulting in $f_\mathrm{SDSS+WISE}=0.97f_\mathrm{s,SDSS+WISE}\left(z_\mathrm{AB}\right)f_\mathrm{p,SDSS+WISE}$. Finally, we obtained a rough estimate of the \citet{2015AA...578A..83G} selection function by comparing the corrected and observed QLFs, i.e.\ taking $f_\mathrm{GOODS-S}=\phi_\mathrm{obs}/\phi_\mathrm{corr}$ (see their Table~3).

@@

\begin{figure*}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{emissivity_18.pdf}
    \includegraphics[width=0.5\textwidth]{emissivity_20.pdf}
    \end{tabular}
  \end{center}
  \caption{LyC emissivity of AGN assuming 100\% escape fraction.
    Model luminosity functions are integrated down to $M_{1450}=-18$
    in the left panel and $-20$ in the right panel.}
  \label{fig:gammapi}
\end{figure*}
